Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-09-21 manuscript/manuscript.tex
Small edits to the Neuquen basin application

\section{Example with real data}

Besides the determination of the distance-size and the delta ratio, we
wanted to illustrate how the algorithm performs under a real data model.
For this task we have chosen the Neuqu\'en Basin: a sedimentary basin
located to the east of the Andes, between 32$^\circ$S and 40$^\circ$S latitude
(see Figure \ref{fig:neuquen-basin}a), that includes continental and marine
siliciclastics, carbonates and evaporites accumulated over the Jurassic and
the Cretaceous constituting a stratigraphic record up to 5000m of depth
\citep{Howell2005}.

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{figures/neuquen-basin.pdf}
\caption{
    Forward model example with real data: Nequ\'en sedimentary modelled
    through tesseroids with exponential density.
    (a) Topography of Neuqu\'en Basin (in km),
    (b) thickness of the sedimentary basin (in meters),
    (c) density function used to model the basin (depth in km and density in
    kg/m$^3$),
    (d)-(l) computed gravity fields: potential $V$ (in J/kg), gradient
    components $g_x$, $g_y$ and $g_z$ (in mGal) and Marussi tensor components
    (in Eötvös), calculated at 50km of height over the ellipsoid.}
\label{fig:neuquen-basin}
\end{figure*}

The thickness of the sediment pack was obtained from \citet{Heine2007}.
We managed to produce a regular grid with a resolution of 0.05$^\circ$ on both
longitude and latitude directions with one value of thickness on each node
(see Figure \ref{fig:neuquen-basin}(b)).

In order to compute the gravity fields generated by this basin we created a
mesh of tesseroids, each one located on one of the nodes of the grid, with
dimensions of 0.05$^\circ$ $\times$ 0.05$^\circ$ and depth equal to the
thickness of the basin on the corresponding node.

We must also define a density function for the entire model.
\citet{Sigismondi2012} measured a minimum and maximum density contrast for
the Neuqu\'en basin of -412kg/m$^3$ and -275kg/m$^3$, respectively.
So we have chosen, for this specific example, an exponential density (like the
one in equation \ref{eq:density-exp}) that assumes the minimum value on the
surface and the maximum at 5858m, i.e. the bottom of the basin, with a value
of $b$ equal to 2km.
This density function can be seen on Figure \ref{fig:neuquen-basin}(c).

We have computed the gravity potential $V$, the gradient components $g_x$,
$g_y$ and $g_z$, and the Marussi tensor components $g_{xx}$, $g_{xy}$,
$g_{xz}$, $g_{yy}$  and $g_{zz}$ on a computation grid of 159$\times$163 nodes
with a 0.05$^\circ$ spacing on both longitude and latitude directions at a
50km height over the reference ellipsoid.
The resulting gravity fields can be seen in Figures
\ref{fig:neuquen-basin}(d)-(l).
We have excluded the other components of the Marussi tensor from the Figure,
although they can be found in the repository.

>>>>

\section{Application to the Neuqu\'en Basin}

We applied the new algorithms and optimal values of $D$ and $\delta$ determined
previously to calculate the gravitational effects of the Neuqu\'en Basin:
a sedimentary basin
located to the east of the Andes, between 32$^\circ$S and 40$^\circ$S latitude
(Fig. \ref{fig:neuquen-basin}a).
The basin includes continental and marine siliciclastics, carbonates, and evaporites
accumulated over the Jurassic and the Cretaceous constituting a stratigraphic record up
to 5000m of depth \citep{Howell2005}.

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{figures/neuquen-basin.pdf}
\caption{
    Gravitational effects of the Nequ\'en sedimentary basin modelled
    using tesseroids with an exponential density function of depth.
    (a) Topography of the Neuqu\'en Basin (in km) and its location in South America,
    (b) thickness of the sedimentary basin \citep[in meters;][]{Heine2007},
    (c) the exponential density function used to model the basin
        (depth in km and density in kg/m$^3$),
    (d)-(l) computed gravitational fields: potential $V$ (in J/kg), gradient
    components $g_x$, $g_y$ and $g_z$ (in mGal), and Marussi tensor components
    (in Eötvös), calculated at 50km of height over the ellipsoid.
}
\label{fig:neuquen-basin}
\end{figure*}

The thickness of the sediment pack was digitized from \citet{Heine2007} on a regular
grid with a resolution of 0.05$^\circ$ on both longitude and latitude directions
(Fig. \ref{fig:neuquen-basin}b).
We created a tesseroid model of the sediment pack by placing a
$0.05^\circ \times 0.05^\circ$ tesseroid on each node of the grid.
The top of each tesseroid was fixed at 0 m depth and the bottom at corresponding
thickness of the basin.

We must also define a density function for the tesseroid model.
\citet{Sigismondi2012} measured a minimum and maximum density contrast for
the Neuqu\'en basin of -412kg/m$^3$ and -275kg/m$^3$, respectively.
We have chosen an exponential density variation (Eq. \ref{eq:density-exp}) that assumes
the minimum value on the top surface and the maximum at 5858m depth (the thickest part
of the basin), with a value of $b$ equal to 2km.
This density function can be seen on Fig. \ref{fig:neuquen-basin}c.

Finally, we computed the gravitational potential $V$, gradient components $g_x$,
$g_y$ and $g_z$, and the Marussi tensor components $g_{xx}$, $g_{xy}$,
$g_{xz}$, $g_{yy}$, and $g_{zz}$ on a computation grid of $159\times163$ nodes
($0.05^\circ$ spacing on both longitude and latitude) at a 50km height over the
reference ellipsoid.
The resulting fields can be seen in Fig.
\ref{fig:neuquen-basin}d-l.
We have excluded the other components of the Marussi tensor from the Fig.
\ref{fig:neuquen-basin},
although they can be found in the online repository.
